### 2018-02-21 paper/sections/1_introduction.tex
paper: camera-ready format; squeeze to 10 pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/cache-management.png}\\
\caption{Using our data management language and policy engine, we design a
dynamically sized caching policy (thick line) for ParSplice.  Compared to
existing configurations (thin lines with \(\times\)'s), our solution saves the most
memory without sacrificing performance and works for a variety of inputs.
Solutions labeled ``no performance degradation" are in comparison to the
baseline of using a cache of unlimited size (``No Cache Management").
\label{fig:cache-management}}
\end{figure}

The biggest challenge for ParSplice is properly sizing the caches in the
storage hierarchy.  The memory usage for a single cache that stores molecule
coordinates is shown in Figure~\ref{fig:cache-management}, where the thin solid
lines marked with \(\times\)'s are the existing configurations in ParSplice.
The default configuration uses an unlimited sized cache, shown by the ``No
Cache Management" line, but using this much memory for one cache is
unacceptable for HPC environments, where a common goal is to keep memory for
such data structures below 3\%\footnote{Anecdotally, this threshold works well
for HPC applications.  For reference, a 1GB cache for a distributed file system
is too large in LANL deployments.}. Furthermore, ParSplice deploys a cache per
300 worker processes, so large simulations need more caches and will use even
more memory.  Users can configure ParSplice to evict data when the cache
reaches a threshold but this solution requires tuning and parameter sweeps; the
``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows how a
poorly configured cache can save memory but at the cost of performance, which
is shown by the text annotation to the right.  Even worse, this threshold
changes with different initial configurations and cluster setups so tuning
needs to be done for all system permutations.  Our dynamically sized cache,
shown by the thick line in Figure~\ref{fig:cache-management}, detects key
access patterns and re-sizes the cache accordingly.  Without tuning or
parameter sweeps, our solution saves more memory than a hand-tuned cache
without any performance degradation, works for a variety of initial conditions,
and could generalize to similar applications. Triggering key eviction at a
certain memory pressure ({\it e.g.} 3\%) requires {\it a priori} system
knowledge while our approach saves the most memory because we model the
behavior of keyspace accesses.

>>>>

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/cache-management.png}\\
\caption{Using our data management language and policy engine, we design a
dynamically sized caching policy (thick line) for ParSplice.  Compared to
existing configurations (thin lines with \(\times\)'s), our solution saves the most
memory without sacrificing performance and works for a variety of inputs.
Solutions labeled ``no performance degradation" are in comparison to the
baseline of a cache of unlimited size.
\label{fig:cache-management}}
\end{figure}

The biggest challenge for ParSplice is properly sizing the caches in the
storage hierarchy.  The memory usage for a single cache that stores molecule
coordinates is shown in Figure~\ref{fig:cache-management}, where the thin solid
lines marked with \(\times\)'s are the existing configurations in ParSplice.
The default configuration uses an unlimited sized cache, shown by the ``No
Cache Management" line, but using this much memory for one cache is
unacceptable for HPC environments, where a common goal is to keep memory for
such data structures below 3\%\footnote{Anecdotally, this threshold works well
for HPC applications.  For reference, a 1GB cache for a distributed file system
is too large in LANL deployments.}. Furthermore, ParSplice deploys a cache per
300 worker processes, so large simulations need more caches and will use even
more memory.  Users can configure ParSplice to evict data when the cache
reaches a threshold but this solution requires tuning and parameter sweeps; the
``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows how a
poorly configured cache can save memory but at the cost of performance, which
is shown by the text annotation to the right.  Even worse, this threshold
changes with different initial configurations and cluster setups so tuning
needs to be done for all system permutations.  Our dynamically sized cache,
shown by the thick line in Figure~\ref{fig:cache-management}, detects key
access patterns and re-sizes the cache accordingly.  Without tuning or
parameter sweeps, our solution saves more memory than a hand-tuned cache
without any performance degradation, works for a variety of initial conditions,
and could generalize to similar applications. Triggering key eviction at a
certain memory pressure ({\it e.g.}, 3\%) requires {\it a priori} system
knowledge while our approach saves the most memory because we model the
behavior of keyspace accesses.

### 2018-02-21 paper/sections/2_parsplice-keyspace-analysis.tex
paper: camera-ready format; squeeze to 10 pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/parsplice.png}\\
\caption{The ParSplice architecture has a storage hierarchy of caches (boxes) and a
dedicated cache process (large box) backed by a persistent database (DB). A splicer
(S) tells workers (W) to generate segments and workers employ tasks (T) for more
parallelization. We focus on the worker's cache (circled), which facilitates
communication and segment exchange between the worker and its tasks.
\label{fig:parsplice}}
\end{figure}

@@

\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keyspace.png}\\

  \caption{The keyspace is small but must satisfy many reads as workers
  calculate segments. Memory usage scales linearly, so it is likely that we will
  need more than one node to manage segment coordinates when we scale the system
  or jobs up. \(\Delta_1\) is the growth rate, an application specific parameter
  described in Section~\S\ref{sec:delta}.

  \label{fig:methodology-keyspace}}
\end{figure}

@@

\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/motivation-regimes.png}\\
  \caption{Key activity for ParSplice starts with many reads to a small set of
keys and progresses to less reads to a larger set of keys.  The line shows the
rate that EOM minima values are retrieved from the key-value store (\(y1\)
axis) and the points along the bottom show the number of unique keys accessed in a 1
second sliding window (\(y2\) axis). Despite having different growth rates
(\(\Delta\)), the structure and behavior of the key activities are similar.
\label{fig:motivation-regimes}}
\end{figure}

@@

\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keys.png}\\
  \caption{Over time, tasks start to access a larger set of keys resulting in
some keys being more popular than others.  Despite different growth rates
(\(\Delta\)), the spatial locality of key accesses is similar between the two
runs.  ({\it e.g.}, some keys are still read 5 times as many times others).
\label{fig:methodology-keys}}
\end{figure}

>>>>

\begin{figure}[t]
\centering
\noindent\includegraphics[width=0.35\textwidth]{figures/parsplice.png}\\
\caption{ParSplice has a storage hierarchy of caches (boxes) and a
dedicated cache process (large box) backed by a persistent database (DB). A splicer
(S) tells workers (W) to generate segments. Workers employ tasks (T) for
parallelization. We focus on the worker's cache (circled), which facilitates
segment exchange.
\label{fig:parsplice}}
\end{figure}

@@

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{.32\linewidth}
    \centering
    \includegraphics[width=1\textwidth]{figures/methodology-keyspace.png}\\
    \caption{Keyspace size and shape.
    \label{fig:methodology-keyspace}}
  \end{subfigure}
  \begin{subfigure}[b]{.32\linewidth}
    \includegraphics[width=0.975\textwidth]{figures/motivation-regimes.png}\\
    \caption{Key accesses over time.
    \label{fig:motivation-regimes}}
  \end{subfigure}
  \begin{subfigure}[b]{.32\linewidth}
    \includegraphics[width=1.05\textwidth]{figures/methodology-keys.png}\\
    \caption{Redundant key accesses.
    \label{fig:methodology-keys}}
  \end{subfigure}
  \caption{Keyspace analysis for ParSplice.
    (a) shows that the keyspace is small but must satisfy many reads. Memory
    usage scales linearly, so we will eventually need more than one node to manage
    segment coordinates for larger jobs. \(\Delta_1\) is the growth rate (see
    Section~\S\ref{sec:delta}).
    (b) shows that key accesses start with many reads to a small set of
    keys and progresses to less reads to a larger set of keys.  The line shows the
    rate that EOM minima values are retrieved from the key-value store (\(y1\));
    points along the bottom show the unique keys accessed in a 1
    second window (\(y2\)). Despite having different
    \(\Delta\)s, the structure of key accesses are similar.
    (c) shows that over time, tasks start accessing a larger set of keys resulting in
    some keys being more popular than others.  Despite different
    \(\Delta\)s, in both runs some keys are \(5\times\) more popular than others.
  }
\end{figure*}

@@

@@

### 2018-02-21 paper/sections/4_dynamic-policies.tex
paper: address reviewer feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: camera-ready format; squeeze to 10 pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% technical details
We implement a basic LRU cache using a ``when" policy of:

\begin{figure}[h]
\footnotesize
\begin{minted}{lua}
server[whoami]['cachesize']>n
\end{minted}
\end{figure}

and a ``how much" policy of:

\begin{figure}[h]
\footnotesize
\begin{minted}{lua}
servers[whoami]['cachesize']-n
\end{minted}
\end{figure}

@@

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
\caption{Policy performance/utilization shows the trade-offs of different sized
caches (\(x\) axis).  ``None" is ParSplice unmodified, ``Fixed Sized Cache"
evicts keys using LRU, and ``Multi-Policy Cache" switches to fixed sized cache
after absorbing the workload's initial burstiness. This parameter sweep
identifies the ``Multi-Policy Cache" of 1K keys as the best solution but this
only works for this system setup and initial configurations.
 \label{fig:methodology-tradeoff}}
\end{figure}

@@

\begin{figure*}[t!]
    \begin{subfigure}[t]{0.48\textwidth}
        \centering
	\includegraphics[width=1\textwidth]{figures/keyspace-zoomed.png}\\

\caption{Key activity for a 4 hour run shows groups of accesses to the same
subset of keys. Detecting these access patterns leads to a more accurate cache
management strategy, which is discussed in Section~\S\ref{sec:regime-detection}
and the results are in Figure~\ref{fig:dscache-vs-none}.
\label{fig:keyspace-zoomed}}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.48\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/dscache-vs-none.png}\\

	\caption{The performance/utilization for the dynamically sized cache (DSCache)
policy. With negligible performance degradation, DSCache adjusts to different initial configurations
(\(\Delta\)'s) and saves 3\(\times\) as much memory in the best case.
\label{fig:dscache-vs-none}}

    \end{subfigure}%
    \caption{Different cache management policies tested over the Mantle policy engine.}
\end{figure*}

@@

The performance and memory utilization is shown by the ``DSCache" bars in
Figure~\ref{fig:dscache-vs-none}. Without sacrificing performance (trajectory
length), the dynamically sized cache policy uses between 32\%-66\% less memory
than the default ParSplice configuration (no cache management) for the 3
initial conditions we test. The memory usage over time is shown by the ``Dynamic Policy"
curves in Figure~\ref{fig:memory-vs-time}, where the behavior resembles the key
access patterns in Figure~\ref{fig:keyspace-zoomed}\footnote{The memory usage
is not {\it exactly} the same because these are two different runs;
Figure~\ref{fig:keyspace-zoomed} has key activity tracing turned on, which
reduces performance.}.  We also show a \(\Delta_2\) growth rate to demonstrate
the dynamic policy's ability to adjust to a different set of initial
conditions.

>>>>

% technical details
We implement an LRU cache using a ``when" policy of:\\
\noindent\texttt{server[whoami]['cachesize']>n}\\
and a ``how much" policy of:\\
\noindent\texttt{servers[whoami]['cachesize']-n}.

@@

\begin{figure*}[t]
  \centering
  \begin{subfigure}[t]{0.32\textwidth}
    \centering
    \includegraphics[width=0.97\textwidth]{figures/methodology-tradeoff.png}\\
    \caption{Policy with system knowledge (\S\ref{sec:arch-specific}).
    \label{fig:methodology-tradeoff}}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\textwidth}
    \centering
    \includegraphics[width=1\textwidth]{figures/keyspace-zoomed.png}\\
    \caption{Key accesses (application knowledge).
    \label{fig:keyspace-zoomed}}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\textwidth}
    \centering
    \includegraphics[width=0.97\textwidth]{figures/dscache-vs-none.png}\\
    \caption{Policy with application knowledge (\S\ref{sec:dom-specific}).
    \label{fig:dscache-vs-none}}
  \end{subfigure}
  \caption{Cache management policies tested over the Mantle policy engine.
    (a) shows performance/utilization trade-offs of different cache
    sizes (\(x\) axis).  ``None" is unmodified, ``Fixed Sized Cache" evicts keys
    using LRU, and ``Multi-Policy Cache" switches to a fixed sized cache after
    absorbing the workload's initial burstiness.  ``Multi-Policy Cache" with 1K
    keys is the best policy but only works for our hardware/configurations.
    (b) shows key accesses for a 4 hour run. Detecting these groups of accesses
    leads to a more accurate cache management strategy.
    (c) shows performance/utilization for the dynamically sized cache
    (DSCache) policy. DSCache adjusts to different initial configurations
    (\(\Delta\)'s) and saves 3\(\times\) as much memory in the best case.
  }
\end{figure*}

@@

@@

The performance and memory utilization is shown by the ``DSCache" bars in
Figure~\ref{fig:dscache-vs-none}. Without sacrificing performance (trajectory
length), the dynamically sized cache policy uses between 32\%-66\% less memory
than the default ParSplice configuration (no cache management) for the 3
initial conditions we test. The memory usage over time is shown by the ``Dynamic Policy"
curves in Figure~\ref{fig:memory-vs-time}, where the behavior resembles the key
access patterns in Figure~\ref{fig:keyspace-zoomed}. The memory usage
is not {\it exactly} the same because these are two different runs;
Figure~\ref{fig:keyspace-zoomed} has key activity tracing turned on, which
reduces performance.  We also show a \(\Delta_2\) growth rate to demonstrate
the dynamic policy's ability to adjust to a different set of initial
conditions.

### 2018-02-21 paper/sections/5_file-systems.tex
paper: address reviewer feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: camera-ready format; squeeze to 10 pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\subsection{Using Cache Management Policies for Load Balancing}

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/trace-atime.png}\\
\caption{File system metadata reads for a Lustre trace collected at LANL. The
vertical lines are the access patterns detected by the ParSplice cache
management policy from Section~\S\ref{sec:dom-specific}. A file system that
load balances metadata across a cluster of servers could use the same pattern
detection to make migration decisions, such as avoiding migration when the
workload is accessing the same subset of keys or keeping groups of accesses
local to a server.  \label{fig:trace-atime}}
\end{figure}

@@

Although other access pattern detection algorithms are possible, we use the one
designed for cache management in Section~\S\ref{sec:regime-detection} with
slight modifications based on our knowledge of file systems\footnote{We
filtered out requests for key IDs less than 2000, as these are most likely path
traversal requests to higher parts of the namespace.}. The vertical lines in
Figure~\ref{fig:trace-atime} are the groups of accesses identified by the
algorithm; it successfully detects the largest group of key accesses that
starts at time 1000 seconds and ends at time 2200 seconds. File systems that
load balance file system metadata across a cluster would want to keep metadata
in that group of key accesses on the same server for locality and would want to
avoid migrating metadata to a different server until the group of key accesses
completes.

@@

% other use cases caches/tiers 
Storage systems have many other data management techniques that would benefit
from the caching policies developed in Sections~\S\ref{sec:arch-specific}
and~\S\ref{sec:dom-specific}. For example, Ceph administrators can use the
policies in ParSplice to automatically size and manage cache
tiers\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/},
caching on object storage devices, or in the distributed block
devices\footnote{http://docs.ceph.com/docs/master/rbd/rbd-config-ref/}.
Integration with Mantle would be straightforward as it is merged into Ceph's
mainline\footnote{http://docs.ceph.com/docs/master/cephfs/mantle/} and the
three caching subsystems mentioned above already maintain key access traces.

@@

  \item Batching: how many operations to group together, when to send large
batches of updates. We could use Mantle to implement pathname leases from
IndexFS~\cite{ren:sc2014-indexfs} or the capabilities from
CephFS\footnote{http://docs.ceph.com/docs/master/cephfs/capabilities/}.

>>>>

\subsection{Using Cache Management Policies for Load Balancing}
\label{sec:using-cache-management-policies-for-load-balancing}
\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/trace-atime.png}\\
\caption{File system metadata reads for a Lustre trace collected at LANL. The
vertical lines are the results of running the pattern detection algorithm
used for cache management in Section~\S\ref{sec:dom-specific}. File system
cases are in
Section~\S\ref{sec:using-cache-management-policies-for-load-balancing}.
\label{fig:trace-atime}}
\end{figure}
%A file system that
%load balances metadata across a cluster of servers could use the same pattern
%detection to make migration decisions, such as avoiding migration when the
%workload is accessing the same subset of keys or keeping groups of accesses
%local to a server.

@@

Although other access pattern detection algorithms are possible, we use the one
designed for cache management in Section~\S\ref{sec:regime-detection} with
slight modifications based on our knowledge of file systems. We
filtered out requests for key IDs less than 2000, as these are most likely path
traversal requests to higher parts of the namespace. The vertical lines in
Figure~\ref{fig:trace-atime} are the groups of accesses identified by the
algorithm; it successfully detects the largest group of key accesses that
starts at time 1000 seconds and ends at time 2200 seconds. File systems that
load balance file system metadata across a cluster would want to keep metadata
in that group of key accesses on the same server for locality and would want to
avoid migrating metadata to a different server until the group of key accesses
completes.

@@

% other use cases caches/tiers 
Storage systems have many other data management techniques that would benefit
from the caching policies developed in Sections~\S\ref{sec:arch-specific}
and~\S\ref{sec:dom-specific}. For example, Ceph administrators can use the
policies in ParSplice to automatically size and manage cache tiers, caching on
object storage devices, or in the distributed block device.  Integration with
Mantle would be straightforward as it is merged into Ceph's mainline and the
three caching subsystems mentioned above already maintain key access traces.

@@

  \item Batching: how many operations to group together, when to send large
batches of updates. We could use Mantle to implement pathname leases from
IndexFS~\cite{ren:sc2014-indexfs} or the capabilities from
CephFS.
